---
output:
  html_document: default
  word_document: default
  pdf_document: default
---

# Sample R Markdown document

```{r}
##load package to import excel file
library(readxl)
##import excel files
data <- read_xlsx("data/cardiac.xlsx", 1)


```

```{r}
## create a scatter plot
library(ggplot2)
data_plot <- ggplot(data, aes(x=Kilograms_body_weight, y=Heart_rate)) + geom_point()
data_plot <- data_plot + ggtitle("Heart rate by body weight") +
            labs(x = "Body weight in Kgs", y = "Heart rate at rest")
data_plot

```

From the above scatter plot we can say that the heart rate at rest and body weight do not have a linear relationship.


Compute $\hat{\beta_0}$ and $\hat{\beta_1}$ and write the regression equation for these data. Interpret the estimated regression coefficients.
```{r}
##compute regression coefficients - linear regression
data_lr <- lm(Heart_rate ~ Kilograms_body_weight, data = data)

##display the summary of the linear regression
summary(data_lr)

```

From the above output we can write  values of *$\hat{\beta_0}$* and *$\hat{\beta_1}$* as below $\hat{\beta_0}$ = 52.42 and $\hat{\beta_1}$ = 0.54 .

Regression equation is as follows:

Heart_rate = $\hat{\beta_0}$ + $\hat{\beta_1}$ * Kilograms_body_weight

Heart_rate = 52.42 + 0.54 * Kilograms_body_weight

Interpretation of coefficients is as follows.

$\hat{\beta_0}$ is 52.42 which means it is the fitted value of heart rate at rest when body weight is 0, in this case this relativeliy meaningless since body weight with 0 weight is not likely to happen.

$\hat{\beta_1}$ is 0.54 which means an increase in body weight by 1 kilogram lead to increase in heart rate at rest by 0.54.


place 95% confidence interval on  $\beta_0$ and $\beta_1$ : we will use SE from the above  regression output.
SE($\hat{\beta_0}$) = 21.1795

SE($\hat{\beta_1}$) = 0.3933

95% confidence interval for $\hat{\beta_0}$ is {$\hat{\beta_0}$ - 1.96 * SE($\hat{\beta_0}$)},
{$\hat{\beta_0}$ + 1.96 * SE($\hat{\beta_0}$)}.

```{r}
upper_interval = 52.4178 + (1.96 * 21.1795)

upper_interval

lower_interval = 52.4178 - (1.96 * 21.1795)

lower_interval
```
Now we can say that the rate of change of conditional mean of heart rate at rest with respect to intercept ($\hat{\beta_0}$) is about 21.1795.

95% confidence interval for $\hat{\beta_1}$ is {$\hat{\beta_1}$ - 1.96 * SE($\hat{\beta_1}$)},
{$\hat{\beta_1}$ + 1.96 * SE($\hat{\beta_1}$)}.

```{r}
upper_interval = 0.5439 + (1.96 * 0.3933)

upper_interval

lower_interval = 0.5439 - (1.96 * 0.3933)

lower_interval

```
Now we can say that the rate of change of conditional mean of heart rate at rest with respect to Kilograms_body_weight ($\hat{\beta_1}$) is about 0.3933.


Plot the regression line on the graph from part 1.1

```{r}

data_plot1 <- ggplot(data, aes(x=Kilograms_body_weight, y=Heart_rate)) + geom_point(color='blue') + geom_smooth(method = "lm", color='red', se = FALSE)
data_plot1 <- data_plot1 + ggtitle("Heart rate by body weight") +
            labs(x = "Body weight in Kgs", y = "Heart rate at rest")
data_plot1
```

Now examine the data point (67, 40). If this data point were removed from the data set, what changes would occur in the estimates of ????0 and ????1?

```{r}
##remove point (67, 40)
data1 <- data[-c(3),]

data1_plot <- ggplot(data1, aes(x=Kilograms_body_weight, y=Heart_rate)) + geom_point(color='blue') + geom_smooth(method = "lm", color='red', se = FALSE)
data1_plot <- data1_plot + ggtitle("Heart rate by body weight") +
            labs(x = "Body weight in Kgs", y = "Heart rate at rest")
data1_plot

```
As we can see from the above scatter plot that, if we remove point (67, 40) regression line goes further away from origin which means coefficient value of intercept $\hat{\beta_0}$ will increase and coefficient value of body weight $\hat{\beta_1}$ will decrease.

```{r}
##regression coefficients
data1_lr <- lm(Heart_rate ~ Kilograms_body_weight, data = data1)

##display the summary of the linear regression
summary(data1_lr)

```
From the above output we can see that the change in the values of intercept $\hat{\beta_0}$ and body weight $\hat{\beta_1}$.
